\bf Visualising How Non-linear Dimension Reduction Warps Your Data

### 2024-01-10 paper.qmd
Add hexbin grid generation with different bins into another script and generated the grids from the source file

library(Rtsne)
library(umap)
library(phateR)
library(reticulate)
library(patchwork)

@@

set.seed(20230531)

@@

tSNE_data <- Fit_tSNE(training_data, opt_perplexity = calculate_effective_perplexity(training_data), with_seed = 20230531)
#(perplexity: ", calculate_effective_perplexity(data), ")
plot_list1 <- plot_tSNE_2D(tSNE_data) + #ggtitle("(a)") + 
  theme_linedraw() +
    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

@@

UMAP_fit <- umap(training_data |> dplyr::select(-ID), n_neighbors = 50, n_components =  2)

UMAP_data <- UMAP_fit$layout |>
  as.data.frame()
names(UMAP_data)[1:(ncol(UMAP_data))] <- paste0(rep("UMAP",(ncol(UMAP_data))), 1:(ncol(UMAP_data)))

UMAP_data <- UMAP_data |>
  mutate(ID = training_data$ID)

@@

#(knn: 5)
# PHATE_data <- Fit_PHATE(training_data, knn = 5, with_seed = 20230531)
# write_csv(PHATE_data, paste0(here::here(), "/data/phate_data_s_curve.csv"))

@@

```{r}
#| echo: false
#| message: false
#| warning: false
 # Loop over loop.vector
  #i <- 1
  num_bins <- 3
  shape_val <- calculate_effective_shape_value(.data = UMAP_data, x = UMAP1, y = UMAP2)

  hexbin_data_object_loop <- extract_hexbin_centroids(UMAP_data, num_bins, shape_val)

  df_bin_centroids_loop <- hexbin_data_object_loop$hexdf_data

  full_centroid_df_loop <- generate_full_grid_centroids(df_bin_centroids_loop)

  full_grid_with_hexbin_id_loop <- dplyr::bind_rows(vec1)[0, ]
  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
    dplyr::mutate_if(is.character, as.numeric)

  for(i in 1:length(sort(unique(full_centroid_df_loop$y)))){

    ## Filter the data set with specific y value
    specific_y_val_df_loop <- full_centroid_df_loop |>
      dplyr::filter(y == sort(unique(full_centroid_df_loop$y))[i])

    ordered_x_df_loop <- specific_y_val_df_loop |>
      dplyr::arrange(x)

    full_grid_with_hexbin_id_loop <- dplyr::bind_rows(full_grid_with_hexbin_id_loop, ordered_x_df_loop)

  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
    dplyr::mutate(hexID = row_number())

  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
    dplyr::rename("c_x" = "x",
           "c_y" = "y")

  full_grid_with_hexbin_id_loop <- dplyr::full_join(full_grid_with_hexbin_id_loop, df_bin_centroids_loop, by = c("hexID" = "hexID")) |>
    dplyr::select(-c(x, y))

  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
      dplyr::mutate(std_counts = counts/max(counts, na.rm = TRUE))

  ## Generate all coordinates of hexagons
  hex_grid_loop <- full_hex_grid(full_centroid_df_loop)

  full_grid_with_polygon_id_df_loop <- map_polygon_id(full_grid_with_hexbin_id_loop, hex_grid_loop)

  full_grid_with_hexbin_id_rep_loop <- full_grid_with_polygon_id_df_loop |>
  dplyr::slice(rep(1:n(), each = 6)) |>
  dplyr::arrange(polygon_id)

  hex_full_count_df_loop <- dplyr::bind_cols(hex_grid_loop, full_grid_with_hexbin_id_rep_loop)

  p1 <-  ggplot(data = hex_full_count_df_loop, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
        geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
    theme_void() +
    theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
          axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
          legend.title = element_text(size=8), #change legend title font size
          legend.text = element_text(size=6)) +
    guides(fill = guide_colourbar(title = "Standardized count")) +
    annotate(geom = 'text', label = "a", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)

@@

```{r}
#| echo: false
#| message: false
#| warning: false
 # Loop over loop.vector
  #i <- 1
  num_bins <- 8
  shape_val <- calculate_effective_shape_value(.data = UMAP_data, x = UMAP1, y = UMAP2)

  hexbin_data_object_loop <- extract_hexbin_centroids(UMAP_data, num_bins, shape_val)

  df_bin_centroids_loop <- hexbin_data_object_loop$hexdf_data

  full_centroid_df_loop <- generate_full_grid_centroids(df_bin_centroids_loop)

  full_grid_with_hexbin_id_loop <- dplyr::bind_rows(vec1)[0, ]
  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
    dplyr::mutate_if(is.character, as.numeric)

  for(i in 1:length(sort(unique(full_centroid_df_loop$y)))){

    ## Filter the data set with specific y value
    specific_y_val_df_loop <- full_centroid_df_loop |>
      dplyr::filter(y == sort(unique(full_centroid_df_loop$y))[i])

    ordered_x_df_loop <- specific_y_val_df_loop |>
      dplyr::arrange(x)

    full_grid_with_hexbin_id_loop <- dplyr::bind_rows(full_grid_with_hexbin_id_loop, ordered_x_df_loop)

  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
    dplyr::mutate(hexID = row_number())

  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
    dplyr::rename("c_x" = "x",
           "c_y" = "y")

  full_grid_with_hexbin_id_loop <- dplyr::full_join(full_grid_with_hexbin_id_loop, df_bin_centroids_loop, by = c("hexID" = "hexID")) |>
    dplyr::select(-c(x, y))

  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
      dplyr::mutate(std_counts = counts/max(counts, na.rm = TRUE))

  ## Generate all coordinates of hexagons
  hex_grid_loop <- full_hex_grid(full_centroid_df_loop)

  full_grid_with_polygon_id_df_loop <- map_polygon_id(full_grid_with_hexbin_id_loop, hex_grid_loop)

  full_grid_with_hexbin_id_rep_loop <- full_grid_with_polygon_id_df_loop |>
  dplyr::slice(rep(1:n(), each = 6)) |>
  dplyr::arrange(polygon_id)

  hex_full_count_df_loop <- dplyr::bind_cols(hex_grid_loop, full_grid_with_hexbin_id_rep_loop)

  p2 <-  ggplot(data = hex_full_count_df_loop, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
    geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
        theme_void() +
    theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
          axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
          legend.title = element_text(size=8), #change legend title font size
          legend.text = element_text(size=6)) +
    guides(fill = guide_colourbar(title = "Standardized count")) +
    annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)

@@

```{r}
#| echo: false
#| message: false
#| warning: false
 # Loop over loop.vector
  #i <- 1
  num_bins <- 15
  shape_val <- calculate_effective_shape_value(.data = UMAP_data, x = UMAP1, y = UMAP2)

  hexbin_data_object_loop <- extract_hexbin_centroids(UMAP_data, num_bins, shape_val)

  df_bin_centroids_loop <- hexbin_data_object_loop$hexdf_data

  full_centroid_df_loop <- generate_full_grid_centroids(df_bin_centroids_loop)

  full_grid_with_hexbin_id_loop <- dplyr::bind_rows(vec1)[0, ]
  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
    dplyr::mutate_if(is.character, as.numeric)

  for(i in 1:length(sort(unique(full_centroid_df_loop$y)))){

    ## Filter the data set with specific y value
    specific_y_val_df_loop <- full_centroid_df_loop |>
      dplyr::filter(y == sort(unique(full_centroid_df_loop$y))[i])

    ordered_x_df_loop <- specific_y_val_df_loop |>
      dplyr::arrange(x)

    full_grid_with_hexbin_id_loop <- dplyr::bind_rows(full_grid_with_hexbin_id_loop, ordered_x_df_loop)

  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
    dplyr::mutate(hexID = row_number())

  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
    dplyr::rename("c_x" = "x",
           "c_y" = "y")

  full_grid_with_hexbin_id_loop <- dplyr::full_join(full_grid_with_hexbin_id_loop, df_bin_centroids_loop, by = c("hexID" = "hexID")) |>
    dplyr::select(-c(x, y))

  full_grid_with_hexbin_id_loop <- full_grid_with_hexbin_id_loop |>
      dplyr::mutate(std_counts = counts/max(counts, na.rm = TRUE))

  ## Generate all coordinates of hexagons
  hex_grid_loop <- full_hex_grid(full_centroid_df_loop)

  full_grid_with_polygon_id_df_loop <- map_polygon_id(full_grid_with_hexbin_id_loop, hex_grid_loop)

  full_grid_with_hexbin_id_rep_loop <- full_grid_with_polygon_id_df_loop |>
  dplyr::slice(rep(1:n(), each = 6)) |>
  dplyr::arrange(polygon_id)

  hex_full_count_df_loop <- dplyr::bind_cols(hex_grid_loop, full_grid_with_hexbin_id_rep_loop)

  p3 <-  ggplot(data = hex_full_count_df_loop, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
        geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
        theme_void() +
    theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
          axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
          legend.title = element_text(size=8), #change legend title font size
          legend.text = element_text(size=6)) +
    guides(fill = guide_colourbar(title = "Standardized count")) +
    annotate(geom = 'text', label = "c", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)

@@

UMAP_fit <- umap(training_data |> dplyr::select(-ID), n_neighbors = 50, n_components =  2)

UMAP_data <- UMAP_fit$layout |>
  as.data.frame()
names(UMAP_data)[1:(ncol(UMAP_data))] <- paste0(rep("UMAP",(ncol(UMAP_data))), 1:(ncol(UMAP_data)))

UMAP_data <- UMAP_data |>
  mutate(ID = training_data$ID)

@@

a1 <- colour_long_edges(.data = distance, benchmark_value = 13, 
                  triangular_object = tr1_object, distance_col = distance) + 
    geom_point(size = 2, colour = "#33a02c") + 
    #ggtitle("(b)") + 
    xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
    theme_light() +
    #coord_equal() +
    theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
          axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
    )

UMAP_data_with_hb_id_selected <- UMAP_data_with_hb_id |> dplyr::filter(hb_id %in% c(14, 177)) |>
  dplyr::mutate(select_point = if_else(hb_id == 14, "data1", "data2"))

a2 <- ggplot(data = hex_full_count_df, aes(x = x, y = y)) + 
  geom_polygon(color = "black", aes(group = polygon_id), fill = "#ffffff") +
  geom_point(data = UMAP_data_with_hb_id_selected, aes(x = UMAP1, y = UMAP2, colour = select_point), alpha = 0.5) +
  scale_color_manual(values=c("#e6550d", "#000000")) +
  geom_point(data = df_bin_centroids_all |> dplyr::filter(hexID %in% c(14, 177)), aes(x = x, y = y), colour = "#33a02c") +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
  theme_light() +
  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
          axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
    )

@@

df1 <- df_all |> 
  dplyr::filter(hb_id == "14") |>
  dplyr::select(tidyselect::starts_with("x")) |>
  dplyr::mutate(type = "data1") ## original dataset

df2 <- df_all |> 
  dplyr::filter(hb_id == "177") |>
  dplyr::select(tidyselect::starts_with("x")) |>
  dplyr::mutate(type = "data2") ## original dataset

df3 <- df_all |> 
  dplyr::filter(!(hb_id %in% c("14", "177"))) |>
  dplyr::select(tidyselect::starts_with("x")) |>
  dplyr::mutate(type = "data") ## original dataset

@@

distance_df_small_edges <- distance |> 
  dplyr::filter(from == 3) |>
  dplyr::filter(to == 49) |>
  head(1)

@@

```{r}
#| echo: false
predict_UMAP_df <- predict(UMAP_fit, test_data |> dplyr::select(-ID)) |>
  as.data.frame()

names(predict_UMAP_df)[1:(ncol(predict_UMAP_df))] <- paste0(rep("UMAP",(ncol(predict_UMAP_df))), 1:(ncol(predict_UMAP_df)))

predict_UMAP_df <- predict_UMAP_df |>
  mutate(ID = test_data$ID)
```

@@

In this section, we demonstrate the effectiveness of our approach through the use of simulation data. These simulations are designed to have known cluster structures and underlying data geometries. We begin with a straightforward example where we generate spherical clusters with a doublet structure. This allows us to assess how well our algorithm can identify and represent these known structures in the low-dimensional space.

@@

```{r}
#| warning: false
#| echo: false 
tSNE_data_gau <- Fit_tSNE(training_data_5 |> dplyr::select(-ID, -cluster), with_seed = 20230531)
```

@@

Screen shots of the **langevitour** with the low-dimensional view applied to tree structured data set, a video of the tour animation is available at (\url{https://drive.google.com/file/d/1YQx4UGlzDify1mGqeDtlqTHncgqidDhq/view}). Combination of the **langevitour** and low-dimensional representation of the tree structured data provides the effectiveness of our approach.
:::

>>>>

library(Rtsne)
library(umap)
library(phateR)
library(reticulate)
library(patchwork)
library(langevitour)

@@

set.seed(20240110)

@@

tSNE_data <- Fit_tSNE(training_data, opt_perplexity = calculate_effective_perplexity(training_data), with_seed = 20240110)
#(perplexity: ", calculate_effective_perplexity(data), ")
plot_list1 <- plot_tSNE_2D(tSNE_data) + #ggtitle("(a)") + 
  theme_linedraw() +
    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

@@

UMAP_data <- read_rds(file = "data/s_curve/s_curve_umap.rds")

@@

#(knn: 5)
# PHATE_data <- Fit_PHATE(training_data, knn = 5, with_seed = 20240110)
# write_csv(PHATE_data, paste0(here::here(), "/data/phate_data_s_curve.csv"))

@@

```{r}
#| echo: false
#| message: false
#| warning: false
hex_full_count_df_loop1 <- read_rds("data/s_curve/s_curve_hex_3.rds")

p1 <-  ggplot(data = hex_full_count_df_loop1, aes(x = x, y = y)) +
      geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
            geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
      scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
        theme_void() +
        theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
              legend.title = element_text(size=8), #change legend title font size
              legend.text = element_text(size=6)) +
        guides(fill = guide_colourbar(title = "Standardized count")) +
        annotate(geom = 'text', label = "a", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)

@@

```{r}
#| echo: false
#| message: false
#| warning: false
hex_full_count_df_loop2 <- read_rds("data/s_curve/s_curve_hex_8.rds")

p2 <-  ggplot(data = hex_full_count_df_loop2, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
    geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
        theme_void() +
    theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
          axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
          legend.title = element_text(size=8), #change legend title font size
          legend.text = element_text(size=6)) +
    guides(fill = guide_colourbar(title = "Standardized count")) +
    annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)

@@

```{r}
#| echo: false
#| message: false
#| warning: false
hex_full_count_df_loop3 <- read_rds("data/s_curve/s_curve_hex_15.rds")

p3 <-  ggplot(data = hex_full_count_df_loop3, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
        geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
        theme_void() +
    theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
          axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
          legend.title = element_text(size=8), #change legend title font size
          legend.text = element_text(size=6)) +
    guides(fill = guide_colourbar(title = "Standardized count")) +
    annotate(geom = 'text', label = "c", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)

@@

# UMAP_fit <- umap(training_data |> dplyr::select(-ID), n_neighbors = 50, n_components =  2)
# 
# UMAP_data <- UMAP_fit$layout |>
#   as.data.frame()
# names(UMAP_data)[1:(ncol(UMAP_data))] <- paste0(rep("UMAP",(ncol(UMAP_data))), 1:(ncol(UMAP_data)))
# 
# UMAP_data <- UMAP_data |>
#   mutate(ID = training_data$ID)

@@

```{r}
#| echo: false
# distance |> dplyr::filter(distance == max(distance))
# df_bin_centroids |> dplyr::filter(row_number() %in% c(10, 65))

a1 <- colour_long_edges(.data = distance, benchmark_value = 15, 
                  triangular_object = tr1_object, distance_col = distance) + 
    geom_point(size = 2, colour = "#33a02c") + 
    #ggtitle("(b)") + 
    xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
    theme_light() +
    #coord_equal() +
    theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
          axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
    )

UMAP_data_with_hb_id_selected <- UMAP_data_with_hb_id |> dplyr::filter(hb_id %in% c(31, 239)) |>
  dplyr::mutate(select_point = if_else(hb_id == 31, "data1", "data2"))

a2 <- ggplot(data = hex_full_count_df, aes(x = x, y = y)) + 
  geom_polygon(color = "black", aes(group = polygon_id), fill = "#ffffff") +
  geom_point(data = UMAP_data_with_hb_id_selected, aes(x = UMAP1, y = UMAP2, colour = select_point), alpha = 0.5) +
  scale_color_manual(values=c("#e6550d", "#000000")) +
  geom_point(data = df_bin_centroids_all |> dplyr::filter(hexID %in% c(31, 239)), aes(x = x, y = y), colour = "#33a02c") +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
  theme_light() +
  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
          axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
    )

@@

df1 <- df_all |> 
  dplyr::filter(hb_id == "31") |>
  dplyr::select(tidyselect::starts_with("x")) |>
  dplyr::mutate(type = "data1") ## original dataset

df2 <- df_all |> 
  dplyr::filter(hb_id == "239") |>
  dplyr::select(tidyselect::starts_with("x")) |>
  dplyr::mutate(type = "data2") ## original dataset

df3 <- df_all |> 
  dplyr::filter(!(hb_id %in% c("31", "239"))) |>
  dplyr::select(tidyselect::starts_with("x")) |>
  dplyr::mutate(type = "data") ## original dataset

@@

distance_df_small_edges <- distance |> 
  dplyr::filter(from == 10) |>
  dplyr::filter(to == 65) |>
  head(1)

@@

```{r}
#| echo: false
predict_UMAP_df <- read_rds("data/s_curve/s_curve_umap_predict.rds")
```

@@

<!--six different clusters-->

<!--Gaussian globe or sphere-->

<!--In this section, we demonstrate the effectiveness of our approach through the use of simulation data. These simulations are designed to have known cluster structures and underlying data geometries. We begin with a straightforward example where we generate spherical clusters with a doublet structure. This allows us to assess how well our algorithm can identify and represent these known structures in the low-dimensional space.

@@

```{r}
#| warning: false
#| echo: false 
tSNE_data_gau <- Fit_tSNE(training_data_5 |> dplyr::select(-ID, -cluster), with_seed = 20240110)
```

@@

Screen shots of the **langevitour** with the low-dimensional view applied to tree structured data set, a video of the tour animation is available at (\url{https://drive.google.com/file/d/1YQx4UGlzDify1mGqeDtlqTHncgqidDhq/view}). Combination of the **langevitour** and low-dimensional representation of the tree structured data provides the effectiveness of our approach.
:::
-->

### 2024-01-10 paper.tex
Add hexbin grid generation with different bins into another script and generated the grids from the source file

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, enhanced]}{\end{tcolorbox}}\fi

@@

In this section, we demonstrate the effectiveness of our approach
through the use of simulation data. These simulations are designed to
have known cluster structures and underlying data geometries. We begin
with a straightforward example where we generate spherical clusters with
a doublet structure. This allows us to assess how well our algorithm can
identify and represent these known structures in the low-dimensional
space.

We then proceed to a more complex example where the data exhibits a more
intricate geometry. This simulation provides a challenging test for our
approach to capture and preserve the underlying data structure
accurately.

We \emph{strongly} recommend viewing the linked videos for each study
while reading. Links to the videos are available in the figures for each
example. The videos show the visual appearance of the
\textbf{langevitour} interface with low-dimensional view and how we can
interact with the tour via the controls.

\textbf{Example 1: Exploring spherical Gaussian clusters with doublet
structure}

The next data set consists of two distinct types of clusters embedded in
a 10-\(d\) space. The first set of clusters includes three Gaussian
clusters with equal variance. Each of these Gaussian clusters is
equidistant from the others, forming a well-separated and symmetrical
arrangement. The second set of clusters consists of three clusters with
a doublet structure. These clusters are also embedded in the 10-\(d\)
space, but each is positioned in the middle of its parent Gaussian
clusters. Moreover, each doublet cluster contains an average number of
points from its corresponding parent Gaussian clusters.

For analysis, we run tSNE on the training data with a perplexity setting
18. Figure~\ref{fig-example3} (a) displays the tSNE results, indicating
that the sub-clusters have been correctly identified. However, the
relative locations of certain clusters to each other appear distorted in
the low-dimensional view, such as the red and pink clusters being far
apart.

Nonetheless, we find no apparent impact on the perception of the data
structure when we examine the tour view and the model (see
Figure~\ref{fig-example3}) through the linked video (see
Figure~\ref{fig-exp1_sc}). The model demonstrates accurate
identification of clusters and captures the internal variations within
each cluster, reflecting the inherent variance present in the data. The
configurations used for the analysis are detailed in
\textbf{?@tbl-table02}. Additionally, the representation of the parent
clusters in the low-dimensional space appears stretched, aligning with
the variance of these clusters. On the other hand, the doublet clusters
are squeezed, reflecting their specific characteristics and relationship
to their parent Gaussian clusters.

{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-example3-1.pdf}

\caption{\label{fig-example3}Visualization and refinement of
low-dimensional representation for sub-clustered data\(\colon\) (a) 2D
layout of tSNE (perplexity = 18), (b) triangular mesh, and (c)
triangular mesh colored by edge type. Edges with lengths less than the
benchmark value of \(2\) are classifiedas small, while edges with
lengths greater than or equal to this value are labeled as long. The
combination of tSNE, hexagonal binning, and triangulation provides the
effectiveness of our approach in visualizing and refining the
low-dimensional representation of the sub-clustered data.}

\includegraphics{figures/example_1a.png}

\subcaption{\label{fig-exp1_sc1}}
\end{minipage}%
%
\begin{minipage}[t]{0.33\linewidth}

\includegraphics{figures/example_1b.png}

\subcaption{\label{fig-exp1_sc2}}
\end{minipage}%
%
\begin{minipage}[t]{0.33\linewidth}

\includegraphics{figures/example_1c.png}

\subcaption{\label{fig-exp1_sc3}}
\end{minipage}%

\caption{\label{fig-exp1_sc}Screen shots of the \textbf{langevitour}
with the low-dimensional view applied to sub-clustered data set, a video
of the tour animation is available at
(\url{https://drive.google.com/file/d/15nAbmc4UoVlUyFcGtnlnmyxrJEbK39nx/view}).
Combination of the \textbf{langevitour} and low-dimensional
representation of the sub-clustered data provides the effectiveness of
our approach.}

\textbf{Example 2: Exploring data with piecewise linear structure}

In our next exploration, we delve into simulated noisy tree-structured
data designed to mimic branching trajectories of cell differentiation.
The simulation is crafted to resemble the hierarchical clustering
pattern observed in cell differentiation trajectories, where mature
cells are akin to the tips of branches.

To begin our analysis, we apply principal components to the data and
focus on the first ten principal components, collectively accounting for
approximately 68\% of the variance explained in the data set. Then, we
run TriMAP with the default parameters on the training data (n\_inliers:
5, n\_outliers: 4, and n\_random: 3). Through this process, we create a
low-dimensional manifold using the configurations detailed in
\textbf{?@tbl-table02}. Finally, we present the model along with the
data using the tour view (Figure Figure~\ref{fig-exp2_sc}), allowing us
to visualize and explore the intricate relationships and structures
present in the data's low-dimensional representation.

{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-example2-1.pdf}

\caption{\label{fig-example2}Visualization and refinement of
low-dimensional representation for tree structured data\(\colon\) (a) 2D
layout of TriMAP (n-inliers = 5, n-outliers = 4, n-random = 3), (b)
triangular mesh, and (c) triangular mesh colored by edge type. Edges
with lengths less than the benchmark value of \(2\) are classified as
small, while edges with lengths greater than or equal to this value are
labeled as long. The combination of TriMAP, hexagonal binning, and
triangulation provides the effectiveness of our approach in visualizing
and refining the low-dimensional representation of the tree structured
data.}

\includegraphics{figures/example_2a.png}

\end{minipage}%
%
\begin{minipage}[t]{0.33\linewidth}

\includegraphics{figures/example_2b.png}

\end{minipage}%
%
\begin{minipage}[t]{0.33\linewidth}

\includegraphics{figures/example_2c.png}

\end{minipage}%

\caption{\label{fig-exp2_sc}Screen shots of the \textbf{langevitour}
with the low-dimensional view applied to tree structured data set, a
video of the tour animation is available at
(\url{https://drive.google.com/file/d/1YQx4UGlzDify1mGqeDtlqTHncgqidDhq/view}).
Combination of the \textbf{langevitour} and low-dimensional
representation of the tree structured data provides the effectiveness of
our approach.}

@@

\hypertarget{zeisel-mouse-brain-strt-seq}{%
\subsection{Zeisel mouse brain
(STRT-Seq)}\label{zeisel-mouse-brain-strt-seq}}

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, interior hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, frame hidden, boxrule=0pt, sharp corners]}{\end{tcolorbox}}\fi

@@

@@

\hypertarget{single-cell-tagged-reverse-transcription-sequencing-data-of-mouse}{%
\subsection{Single-Cell Tagged Reverse Transcription sequencing data of
mouse}\label{single-cell-tagged-reverse-transcription-sequencing-data-of-mouse}}
